---
title: "Extracting NIBRS Data for the Network Analysis of Clearance Crime Rates Across the Country
"
author: "Mahlika George"
output: pdf_document
---
## Storing Variables 

To start the process of exporting the data into a formulated excel sheet, I first read in the data from the stored csv. The variables are than stored accordingly. 

```{r}
Data <- read.csv("C:/Users/mkg4541/Downloads/Sorted.csv", header = TRUE)
City <- Data$BH007
State <- Data$BH008
cities <- list()

#The following loop goes through all of the entries of cities listed in NIBRS and adds them to a new list only if they have not been added already. This is to remove duplicates.
for (i in City) {
    if (length(cities)==0) {
      cities <- append(cities,i)
    }
    else if (!i %in% cities) {
      cities <- append(cities,i)
    }
}

#The following code resaves city to only store the top 100 cities by population
cities <- head(cities,-(length(cities)-100))

#The following stores the index in where the last city first appears in NIBRS so we can find the associated state
#This loop finds where that city and state combo are last used in NIBRS so we can create a shortened data frame
index <- which(City %in% cities[100])[1]
s <- State[index]
for (i in which(City %in% cities[100])) {
  if(State[i] == s) {
    iny <- i
  }
  else{
    break
  }
}

#A new data fram is created for only the top 100 cities
new_Data <- Data[1:iny,]

```

## Calculating the Incidents and Rates

Next, I then created several lists to calculate and store the total number of incidents of certain crimes and all crimes. That data is calculated with a series of loops and conditionals. 

```{r}
#Several new variables are created for future storing
Counted <- list()

#Total Incidents
sum = list()

AA <- list()
Murder <- list()
Robbery <- list()
sum[1] = 0
AA[1] <- 0
Murder[1] <- 0
Robbery[1] = 0

#This variable is used to iterate through the list
index <- 1

#For numbers between 1 and the last index of the new data frame
for (i in 1:iny) {
  
  #if the city has not yet been added to the list "Counted", and the sum is now 1
  if (!as.character(new_Data$BH007[i]) %in% Counted) {
    Counted[index] <- as.character(new_Data$BH007[i])
    sum[index] <- 1
    
    #if the offense code is 120, a robbery was the type of crime reported and the robbery list at that index is 1
     if (as.character(new_Data$V20061[i]) =="120") {
       Robbery[index] <- 1
     }
    else {
      #otherwise, if it is not a robbery, 0 robberies have been reported and thus 0
      Robbery[index] <- 0
    }
    #if the other crime offense was a robbery, add 1 to the current total
    if (as.character(new_Data$V20062[i]) =="120") {
        Robbery[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Robbery[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #if the other crime offense was a robbery, add 1 to the current total
    if (as.character(new_Data$V20063[i]) =="120") {
      Robbery[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Robbery[match(as.character(new_Data$BH007[i]),Counted)])
    }
    
    #if the offense code is 91, a murder was the type of crime reported and the murder list at that index is 1
    if (as.character(new_Data$V20061[i]) =="91") {
      Murder[index] <- 1
    }
    
    #otherwise, if it is not a murder, 0 murders have been reported and thus 0
    else {
      Murder[index] <- 0
    }
    
    #if the other crime offense was a murder, add 1 to the current total
    if (as.character(new_Data$V20062[i]) =="91") {
      Murder[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Murder[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #if the other crime offense was a murder, add 1 to the current total
    if (as.character(new_Data$V20063[i]) =="91") {
      Murder[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Murder[match(as.character(new_Data$BH007[i]),Counted)])
    }
    
    #if the offense code is 131, an assult was the type of crime reported and the assault list at that index is 1
    if (as.character(new_Data$V20061[i]) =="131") {
      AA[index] <- 1
    }
   
    #otherwise, if it is not an assault, 0 assaults have been reported and thus 0
    else {
      AA[index] <- 0
    }
    #if the other crime offense was an assault, add 1 to the current total
    if (as.character(new_Data$V20062[i]) =="131") {
      AA[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(AA[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #if the other crime offense was an assault, add 1 to the current total
    if (as.character(new_Data$V20063[i]) =="131") {
      AA[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(AA[match(as.character(new_Data$BH007[i]),Counted)])
    }
    
    #we now move to sorting in the next position in the new lists
    index <- index + 1
   
  }
  else {
    #since that city was seen again, ad not for the first time, add 1 to its current incident total
    sum[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(sum[match(as.character(new_Data$BH007[i]),Counted)])
    
    #since that city already has a stored number for the robbery total, add 1 to the robbery total
    if (as.character(new_Data$V20061[i]) =="120") {
        Robbery[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Robbery[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #since that city already has a stored number for the robbery total, add 1 to the robbery total
    if (as.character(new_Data$V20062[i]) =="120") {
        Robbery[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Robbery[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #since that city already has a stored number for the robbery total, add 1 to the robbery total
    if (as.character(new_Data$V20063[i]) =="120") {
      Robbery[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Robbery[match(as.character(new_Data$BH007[i]),Counted)])
    }
    
    #since that city already has a stored number for the murder total, add 1 to the murder total
    if (as.character(new_Data$V20061[i]) =="91") {
        Murder[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Murder[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #since that city already has a stored number for the murder total, add 1 to the murder total
    if (as.character(new_Data$V20062[i]) =="91") {
        Murder[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Murder[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #since that city already has a stored number for the murder total, add 1 to the murder total
    if (as.character(new_Data$V20063[i]) =="91") {
      Murder[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(Murder[match(as.character(new_Data$BH007[i]),Counted)])
    }
    
    #since that city already has a stored number for the assault total, add 1 to the assault total
    if (as.character(new_Data$V20061[i]) =="131") {
        AA[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(AA[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #since that city already has a stored number for the assault total, add 1 to the assault total
    if (as.character(new_Data$V20062[i]) =="131") {
        AA[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(AA[match(as.character(new_Data$BH007[i]),Counted)])
    }
    #since that city already has a stored number for the assault total, add 1 to the assault total
    if (as.character(new_Data$V20063[i]) =="131") {
      AA[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(AA[match(as.character(new_Data$BH007[i]),Counted)])
    }
  }
   

}
#Store the final city list as the Counted list
final_city <- Counted
final_state <- list()
final_population <- list()
j <-1

#For every city in the Counted list, add there corresponding State from the data frame, in the same matching index but within a new list "final_state"
for (i in final_city) {
  final_state[j] <- as.character(new_Data$BH008[match(i,new_Data$BH007)])
  final_population[j] <- as.character(new_Data$BH019[match(i,new_Data$BH007)])
  j <- j + 1
  }

```

Now that all of the values have been calculated, write it out to a new excel sheet

```{r}
#Download needed libraries
library("xlsx")

#Creating a new data frame of all of the new variables, and unlisting them so they are in the right format to do so
tes <- data.frame("Total Population" = matrix(unlist(final_population)), "City" = matrix(unlist(final_city)), "State" = matrix(unlist(final_state)), "Total Incidents by city" = matrix(unlist(sum)), "Aggravated Assault Incidents" = matrix(unlist(AA)), "Murder/Nonnegligent Manslaughter Incidents" = matrix(unlist(Murder)), "Robbery Incidents" = matrix(unlist(Robbery)))

#Writing the new data frame out to excel 
write.xlsx2(tes, file = "testing.xlsx", col.names = TRUE, row.names = TRUE)
```

